Rice yield improvements through plant breeding are offset by inherent yield declines over time

### 2018-03-18 output/breed_paper_revised.Rnw
Bruce's final edits?

Based on Model 3, there was no evidence of consistent differences between cultivars in
the rate of yield decline over time (i.e., all cultivar specific rates
of decline overlapped with eachother and zero). Differences between cultivars in the decline
of yield over time, seen in the raw data (Fig. \ref{fig:all_cult}),
are therefore likely small and uncertain. Given this, the model excluding varying
rates of decline by cultivar (Model 2) is presented here. The yields
of each cultivar over time were evaluated relative to when the
cultivar was released using the full model which accounts for both
yield decline over time and yield improvements over time (Model 2). The average decrease
in yield over time following a cultivar's release
was estimated at \Sexpr{yrs_coef} kg ha\tsups{-1} year\tsups{-1} (90\%
CI: \Sexpr{RES_CI[2]} to \Sexpr{RES_CI[1]}) (Fig. \ref{fig:RES_year};
Table \ref{tbl:coef}). The oldest cultivar in the trials
(\Sexpr{gsub("M","M-",unique(mod_data[which.min(mod_data[,"release_yr"]),"id"]))})
is estimated to have lost an average of
\Sexpr{round(yrs_coef*-36/1000,1)} Mg ha\tsups{-1} since its
release. Likewise, an estimated 560 kg ha\tsups{-1} yield
advantage of M\=/206 over M\=/202 (the current and former predominant
cultivars, respectively), can be attributed solely to the decline
of M\=/202's yield over time (Figs. \ref{fig:first_yrs} \&
\ref{fig:compare_yrs}). Further supporting this estimate of yield
decline, both M\=/202 and M\=/205 yielded similarly at their respective year of
release (Fig. \ref{fig:compare_yrs}), despite the fact that M\=/202 was
released 19 years prior to M\=/206 (Fig. \ref{fig:compare_yrs}; Table
\ref{tbl:ids}).

@@

Our investigation here is narrowly focused on the rough-rice/field
grain yield. Other advancements due to plant improvement, including
increased quality~\citep{mckenzie1993} and milling
yield~\citep{mckenzie1994}, have been made over the study period and
are important. Additionally, crop duration has been reduced in CA rice
cultivars. For example, while the predominate cultivars M\=/202 and
M\=/206 had similar yield potentials at the time of release (Fig. 5),
M\=/206 matures about 6 days earlier than M\=/202 (Table
\ref{tbl:ids}). Shorter duration is very important in CA due to an
already short growing season and water limitations (shorter duration
cultivars would presumably require less irrigation water). For a
detailed discussion of cultivars' agronomic and grain quality
characteristics, refer to \url{http://rice.ucanr.edu/Reports-Publications/Rice_Production_Workshop_Manual/}.

>>>>

Based on Model 3, there was no evidence of differences between cultivars in
the rate of yield decline over time (i.e., all cultivar specific rates
of decline overlapped with eachother). Differences between cultivars in the decline
of yield over time, seen in the raw data (Fig. \ref{fig:all_cult}),
are therefore likely small and uncertain. Given this, the model excluding varying
rates of decline by cultivar (Model 2) is presented here. The yields
of each cultivar over time were evaluated relative to when the
cultivar was released using the full model which accounts for both
yield decline over time and yield improvements over time (Model 2). The average decrease
in yield over time following a cultivar's release
was estimated at \Sexpr{yrs_coef} kg ha\tsups{-1} year\tsups{-1} (90\%
CI: \Sexpr{RES_CI[2]} to \Sexpr{RES_CI[1]}) (Fig. \ref{fig:RES_year};
Table \ref{tbl:coef}). The oldest cultivar in the trials
(\Sexpr{gsub("M","M-",unique(mod_data[which.min(mod_data[,"release_yr"]),"id"]))})
is estimated to have lost an average of
\Sexpr{round(yrs_coef*-36/1000,1)} Mg ha\tsups{-1} since its
release. Likewise, an estimated 560 kg ha\tsups{-1} yield
advantage of M\=/206 over M\=/202 (the current and former predominant
cultivars, respectively), can be attributed solely to the decline
of M\=/202's yield over time (Figs. \ref{fig:first_yrs} \&
\ref{fig:compare_yrs}). Further supporting this estimate of yield
decline, both M\=/202 and M\=/205 yielded similarly at their respective year of
release (Fig. \ref{fig:compare_yrs}), despite the fact that M\=/202 was
released 19 years prior to M\=/206 (Fig. \ref{fig:compare_yrs}; Table
\ref{tbl:ids}).

@@

Our investigation here is narrowly focused on the rough-rice/field
grain yield. Other advancements due to plant improvement, including
increased quality~\citep{mckenzie1993} and milling
yield~\citep{mckenzie1994}, have been made over the study period and
are important. Furthermore, as previously mentioned, important disease
resistance traits have been developed. Additionally, crop duration has been reduced in CA rice
cultivars. For example, while the predominate cultivars M\=/202 and
M\=/206 had similar yield potentials at the time of release (Fig. 5),
M\=/206 matures about 6 days earlier than M\=/202 (Table
\ref{tbl:ids}). Shorter duration is very important in CA due to an
already short growing season and water limitations (shorter duration
cultivars would presumably require less irrigation water). For a
detailed discussion of cultivars' agronomic and grain quality
characteristics, refer to \url{http://rice.ucanr.edu/Reports-Publications/Rice_Production_Workshop_Manual/}.
